Co-moving stars in \textsl{Gaia DR1}

### 2016-11-17 paper/ms.tex
Edit abstract and text (from DNS)

\begin{abstract}
% Context
The primary sample of the \gaia\ \textsl{First Data Release} is the brand-new
\textsl{Tycho-Gaia Astrometric Solution}. The precision and size of this sample
creates an opportunity to find new binary stars and moving groups, especially
rare binaries or sparsely populated groups.
% Aims
Here we seize this opportunity.
% Methods
We use a justified marginalized likelihood ratio test to separate
pairs of stars with surprisingly similar three-space velocities from
those consistent with being drawn independently from the field
population.  Although we perform some visualizations using a
(bias-corrected) inverse parallax as a distance, the likelihood test
works in the observable space and uses the \gaia\ noise model
responsibly.
% Results
We find pairs of co-moving stars out to very wide
separations up to 10~pc, the limit of our search.
Some of these pairs form larger networks through mutual co-moving neighbors,
many of which correspond to the known open clusters, and stars in OB associations.
However, there still is a huge number of very wide ($>3$~pc) separation pairs of
stars that are co-moving.
We discuss their nature, and the prospects for testing stellar models and chemical
abundance measurements using our catalog.

% change to probabilistic statement "co-moving"
% we present a table of co-moving pairs
% add dynamical inference
% the last line = the most important thing

@@

At small separations ($0.001-1$~pc), they correspond to wide binaries (and
multiples) that are either weakly gravitationally bound or currently being
dispersed.
Because they have low binding energies, a sample of wide binaries is valuable
for investigating the Galactic dynamical environment.
These systems must have survived their dynamic birth environment which
provides a window to the star formation process.
They may also be disrupted by Galactic tides, or other massive objects
such as other stars, molecular clouds, or MACHOs \citep{Yoo:2004aa,Allen:2014aa}.
Wide binaries are also good testbeds for stellar evolution models.
For example, main sequence-red giant wide binaries are especially valuable in
testing consistency between stellar atmosphere models of these two types as the
two stars were likely born at the same time with the same chemical compositions,
but evolved indepdently because of their wide separation.
Finally, calibration of stellar parameters of low-mass stars (e.g., M dwarfs),
which dominate the stellar content of the Galaxy by number, can benefit from
a larger sample of widely separated binaries containing a low mass star and a
much brighter FGK star whose stellar parameters are easier to measure
\citep[e.g.,][]{Rojas-Ayala:2012aa}.

@@

The abundance of pairs of stars with small velocity difference in
\figname~\ref{fig:dv-sep} suggests that there are a
significant number of co-moving stars in the \tgas\ data at a range
of separations.
Here we develop a method to select high-confidence co-moving
stars that properly incorporates the complex uncertainties associated with the
\gaia\ data. We make the following assumptions in order to construct a
statistical model (a likelihood function with explicit priors on our
parameters):
\begin{itemize}
  \item We assume that the uncertainties in the data---parallax, $\varpi$, and
    two proper motion components, $\propm = (\begin{array}[t]{c c} \mu_\alpha &
    \mu_\delta\end{array})^\mathsf{T}$---are Gaussian with known covariances
    $\mat{C}$. The values of covariances are provided by the \gaia\ data
    \citep{Lindegren:2012aa,Lindegren:2016aa}.
  \item We assume that the 3-space velocities of a given pair of stars
    $(\vec{v}_i, \vec{v}_j)$ in the \tgas\ sample (relative to the solar system
    barycenter) are either (1) co-moving with velocity drawn from a velocity
    prior $p(\vec{v})$ and  velocity difference drawn from a zero-mean Gaussian
    with velocity dispersion $s$, or (2) individually drawn from the same
    velocity prior $p(\vec{v})$.
\end{itemize}

@@

We show the distribution co-moving pairs in galactic longitude and
distance in Figure~\ref{fig:glon_d_pairlines}.
The connection between known co-moving structures and the networks of co-moving
pairs found in this work becomes immediately clear when
we overplot the positions of known Milky Way star clusters \citep{Kharchenko:2016aa},
and stars in OB associations \citep{de-Zeeuw:1999aa}.
Many of the larger networks are clearly associated with known star clusters.
The most prominent is Melotte 111 at
$(l,d)\approx (220\arcdeg, 87~\textrm{pc})$ otherwise known as the Coma star cluster.
Clumps of co-moving pairs at $(l,d)\approx (300-360\arcdeg, 100-200~\textrm{pc})$
seem to strongly correlate with the locations of OB associations
Upper Scorpius, Upper Centaurus Lupus, and Lower Centaurus Crux
\citep{de-Zeeuw:1999aa}.

>>>>

\begin{abstract}
% Context
The primary sample of the \gaia\ \textsl{First Data Release} is the brand-new
\textsl{Tycho-Gaia Astrometric Solution} (\tgas). The precision and size of this sample
presents an opportunity to find new binary stars and moving groups, especially
rare binaries or sparsely populated groups.
% Aims
Here, we look for high-confidence co-moving pairs of stars using this data.
% Methods
We use a justified marginalized likelihood ratio test to separate
pairs of stars with similar three-space velocities from
those consistent with being drawn independently from the field
population.  Although we perform some visualizations using a
(bias-corrected) inverse parallax as a distance, the likelihood test
works in the observable space and uses the \gaia\ noise model
responsibly.
% Results
We find pairs of co-moving stars out to very wide
separations up to 10~pc, the limit of our search.
Some of these pairs form larger networks through mutual co-moving neighbors,
many of which correspond to the known open clusters, and stars in OB associations.
However, there still is a surprisingly large number (\todo{SMOH: give number})
of very wide ($>3$~pc) separation pairs of stars that are co-moving.
We present a catalog of high-confidence co-moving pairs of stars in \tgas\
with auxiliary information,
and discuss its potential usage
for making dynamical inferences, and testing stellar models and chemical
abundance measurements.

@@

At small separations ($0.001-1$~pc), they are wide binaries (and
multiples) that are either weakly gravitationally bound or
slowly separating.
Because they have low binding energies, a sample of wide binaries is valuable
for investigating the Galactic dynamical environment.
These systems must have survived both their dynamic birth environment and 
have avoided tidal destruction along their orbit.
Thus, the statistical properties of wide binaries provide a window into both the star formation
process and the Galactic dynamics
including Galactic tides, and other massive perturbers such as molecular clouds
and MACHOs \citep{Yoo:2004aa,Allen:2014aa}.
Wide binaries are also good testbeds for stellar evolution models.
For example, main sequence-red giant wide binaries are especially valuable in
testing consistency between stellar atmosphere models as the
two stars were likely born at the same time with the same chemical compositions,
but evolved indepdently because of their wide separation.
Finally, calibration of stellar parameters of low-mass stars (e.g., M dwarfs),
which dominate the stellar content of the Galaxy by number, can benefit from
a larger sample of widely separated binaries containing a low mass star and a
much brighter FGK star whose stellar parameters are easier to measure
\citep[e.g.,][]{Rojas-Ayala:2012aa}.

@@

The abundance of pairs of stars with small velocity difference in
\figname~\ref{fig:dv-sep} suggests that there are a
significant number of co-moving stars in the \tgas\ data at a range
of separations.
Here we develop a method to select high-confidence co-moving
stars that properly incorporates the complex uncertainties associated with the
\gaia\ data. We make the following assumptions in order to construct a
statistical model (a likelihood function with explicit priors on our
parameters):
\begin{itemize}
  \item We assume that the uncertainties in the data---parallax, $\varpi$, and
    two proper motion components, $\propm = (\begin{array}[t]{c c} \mu_\alpha &
    \mu_\delta\end{array})^\mathsf{T}$---are Gaussian with known covariances
    $\mat{C}$. The values of covariances are provided as part of the \gaia\ data
    \citep{Lindegren:2012aa,Lindegren:2016aa}.
  \item We assume that the 3-space velocities of a given pair of stars
    $(\vec{v}_i, \vec{v}_j)$ in the \tgas\ sample (relative to the solar system
    barycenter) are either (1) co-moving with velocity drawn from a velocity
    prior $p(\vec{v})$ and  velocity difference drawn from a zero-mean Gaussian
    with velocity dispersion $s$, or (2) individually drawn from the same
    velocity prior $p(\vec{v})$.
\end{itemize}

@@

We show the distribution of co-moving pairs in galactic longitude and
distance in Figure~\ref{fig:glon_d_pairlines}.
The connection between known co-moving structures and the networks of co-moving
pairs found in this work becomes immediately clear when
we overplot the positions of known Milky Way star clusters \citep{Kharchenko:2016aa},
and stars in OB associations \citep{de-Zeeuw:1999aa}.
Many of the larger networks are clearly associated with known star clusters.
The most prominent is Melotte 111 at
$(l,d)\approx (220\arcdeg, 87~\textrm{pc})$ otherwise known as the Coma star cluster.
Clumps of co-moving pairs at $(l,d)\approx (300-360\arcdeg, 100-200~\textrm{pc})$
seem to strongly correlate with the locations of OB associations
Upper Scorpius, Upper Centaurus Lupus, and Lower Centaurus Crux
\citep{de-Zeeuw:1999aa}.
